\section{Introduction}
\textbf{The Learning Problem of Pattern Recognition}
\begin{itemize}
    \item[-]Representation of Objects: Data Representation.
    \item[-]Definition/Modeling of structure: What is a pattern?
    \item[-]Optimization: Search for preferred structures.
    \item[-]Validation: are the structures indeed in the data or are they explained by fluctuations?
\end{itemize}{}
\\
\textbf{Feature Space}
\begin{itemize}
    \item[-]Measurement space X: mathematical space in which data is represented (e.g. numerical R, Boolean, Categorical)
    \item[-]Features: derived quantities (edges, corners, motion vectors,...)
    \item[]Remark: The selection of a specific feature space predetermines the metric to compare data.
\end{itemize}
\\
\textbf{Typical learning problems}
\begin{itemize}
    \item[-]Classification: learning an indicator function.
    \item[-]Regression: learning a real valued function.
    \item[-]Density estimation: learning a probability density of the data source.
    \item[-]Dimension reduction: learning a linear or nonlinear projection.
    \item[-]Data compression: learning a coding efficient representation.
\end{itemize}{}
\\
\textbf{Key Concepts of Machine Learning:}
\begin{itemize}
    \item[-]Trade off training error vs. model complexity
    \begin{itemize}
        \item[-]Regularizers prevent overfitting.
        \item[-]Hyper parameter/model selection via cross-validation.
    \end{itemize}{}
    \item[-]Corresponodence prob. modeling: Loss = Likelihood, regularizer = prior
    \item[-]Kernel trick: replace inner products by kernel function.
    \item[-]Neural nets for feature learning.
    \item[-]Discriminative vs. generative models
    \begin{itemize}
        \item[$\rightarrow$]Discrim.: Learn function $f: X \rightarrow Y$
        \item[$\rightarrow$]Genr.: Learn joint distribution $P(X=x, Y=y)$
    \end{itemize}{}
    \item[-]Unsupervised learning as latent variable modeling (clustering = classification, dim. Reduction = regression)
    $\rightarrow$ Training by EM algorithm.
\end{itemize}{}
\\